---
title: "Global Sewage Project"
---

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width=10, fig.height=5)
```

```{r, message=FALSE, warning=FALSE}
library(vegan)
library(tidyverse)
library(MASS)
library(GGally)
library(betareg)
library(lmtest)
library(pheatmap)
library(patchwork)
```

# Short introduction 
The main objective of this project is to use untreated sewage as a marker for antibiotic resistance prevalence in clinical pathogens.  
We want to test whether sewage meatgenomic could be used to predict clinical resistance. 

## The data
Metagenomic sequencing data from untreated sewage comes from an article in *Nature Communications* [Global monitoring of antimicrobial resistance based on metagenomics analyses of urban sewage.](https://doi.org/10.1038/s41467-019-08853-3)  
The raw sequence data from this study was downloaded from [ENA](https://www.ebi.ac.uk/ena) under project [ERP015409](https://www.ebi.ac.uk/ena/data/view/PRJEB13831).

The data consists of 234 sewage metagenomemes from 62 countries. The antibiotic resistance genes were annotated against [ResFinder v.3.1.0](https://bitbucket.org/genomicepidemiology/resfinder) database and *intI1* integrase gene against [MobileGeneticElementDatabase](https://github.com/KatariinaParnanen/MobileGeneticElementDatabase).  

The clinical resistance data for *E. coli* against four different classes of antibiotics, aminopenicillins, fluoroquinolones, 3rd generation cephalosporins and aminoglycosides was collected from different surveillance networks. Only data from more than 100 isolates was considered reliable and included in the actual analyses.  
In addition, aggregated resistance index from  [Collignon *et al.*, 2018](https://doi.org/10.1016/S2542-5196(18)30186-4) was used in the study.

Socio-economical factors were retrieved from [The World Bank Databank](https://databank.worldbank.org/source/health-nutrition-and-population-statistics).

All ARG and *intI1* integrase gene counts were normalized to the amount of sequence data (million bases). Since the clinical resistance data is on country level, mean of the normalized gene count data was taken for countries with more than one sewage sample. 

All data has been gathered to one data frame and is available in the `Data` folder. 

# Data Analyses 
Read in the data and have a look. 
```{r, echo=FALSE}
# Data
load("../Data/GlobalSewage.R", verbose=TRUE)
glimpse(ResData)
```

## Main article

### Figure 1 and 

## Supplementary materials

### Suppl. Fig 1. 